Field-informed Reinforcement Learning for Learning Large-Scale Collective Tasks

### 2023-05-19 paper.tex
feat: on motivation incipit

\begin{equation}
h_{v}^{(k)}=\operatorname{COMBINE}^{(k)}\left(h_{v}^{(k-1)}, a_{v}^{(k)}\right)
\end{equation}
where $h_{v}^{k}$ is the embedding of node $v$ within the $k$-th layer, 
 $\mathcal{N}_G(v)$ is the set of neighbours of node $v$ computed from $E$.
%
$h_v{v}^0$ is the initial embedding of node $v$ and it is usually set to the node's feature vector $f_v$.
%
The differential part comes into play in the $\operatorname{COMBINE}$ function, 
 which is usually a differentiable function such as a neural network layer.
%
$\operatorname{AGGREGATE} $ instead is a function that aggregates the information from the neighbours of a node $v$
 and it could be a simple sum, max or sum of products. 
 More complex aggregation functions are available~\cite{pellegrini2020learning}.
%
Thereafter, we express the application of a \ac{GNN} to a graph $G_f$ as:
\begin{equation}
GNN(G_f) = \{h_v^{(k)}: v \in V, k \in \mathbb{N}\}
\end{equation} 
This formulation allows \acp{GNN} to effectively process and extract features from graph-structured data by iteratively aggregating and transforming information from the node's neighbours.
%\lukas{do we need definitions of $o_v$, $h_v^*$ over time $t$?}

@@

This model can be used then to express the evolution of the system in time.
Specifically, starting from a global state $\mathcal{S}^P_t$, the next state $\mathcal{S}^P_{t+1}$ is defined as:
\begin{equation}
\mathcal{A}^P_t = \pi(\xi(S^P_t))
\end{equation}
\begin{equation}
\mathcal{S}^P_{t+1} = \mathcal{T}(\mathcal{S}^P_t, \mathcal{A}^P_t)
\end{equation}
Given a time $t$, the system can be also represented as a graph $G_t = (V_t, E_t)$, 
  where $E_t$ is built from $\mathcal{N}$.
  Each node is then decorated with the local observation perceived at the time $t$: $o^v_t \in O$. 
  Therefore, this graph can be used for being processed by a \ac{GNN} as done in~\cite{DBLP:conf/corl/TolstayaGPP0R19,tolstaya2020learning,DBLP:conf/icra/GosrichMLPYR022} if we set $f_v = o^v_t$.
\subsection{Motivation}
The definition of distributed controllers for intelligent systems can be divided into two major approaches: 
 \emph{manual design}, where the application is conceived and described through a programming language, 
 and \emph{automatic design}~\cite{brambilla2013swarm}, which employs machine learning techniques to synthesise distributed controllers. 
%
Our work lies between these two approaches, specifically in the areas of field-based coordination and \ac{MAARL} with the use of \acp{GNN}.

Compared to earlier works on field-based coordination, 
 our approach builds on the concept of co-fields~\cite{DBLP:journals/pervasive/MameiZL04}, 
 where agents exploit the field as a stigmergic medium to receive system-wide information and apply reasoning to this data.
%  
We present a subsequent approach where agent intelligence is synthesised through \ac{MAARL} and the use of \ac{GNN} to learn a local representation from the neighbourhood.
%
By learning a smart policy directly in the environment, the agent becomes capable of adapting its behaviour to new situations.

@@

As value-based algorithm we relied on \ac{DQN}~\cite{mnih2015playing} \ga{to introduce in many rl??} with two major modifications (see \Cref{alg:dqn}): 

>>>>

\begin{equation}
h_{v}^{(k)}=\operatorname{COMBINE}^{(k)}\left(h_{v}^{(k-1)}, a_{v}^{(k)}\right)
\end{equation}
where $h_{v}^{k}$ is the embedding of node $v$ within the $k$-th layer, 
 $\mathcal{N}_G(v)$ is the set of neighbours of node $v$ computed from $E$.
%
$h_v^0$ is the initial embedding of node $v$ and it is usually set to the node's feature vector $f_v$.
%
The differential part comes into play in the $\operatorname{COMBINE}$ function, 
 which is usually a differentiable function such as a neural network layer.
%
$\operatorname{AGGREGATE} $ instead is a function that aggregates the information from the neighbours of a node $v$
 and it could be a simple sum, max or sum of products. 
 More complex aggregation functions are available~\cite{pellegrini2020learning}.
%
Thereafter, we express the application of a \ac{GNN} to a graph $G_f$ as:
\begin{equation}
GNN(G_f) = \{h_v^{(k)}: v \in V, k \in \mathbb{N}\}
\end{equation} 
This formulation allows \acp{GNN} to effectively process and extract features from graph-structured data by iteratively aggregating and transforming information from the node's neighbours.
%\lukas{do we need definitions of $o_v$, $h_v^*$ over time $t$?}

@@

This model can be used then to express the evolution of the system in time.
Specifically, starting from a global state $\mathcal{S}^P_t$, the next state $\mathcal{S}^P_{t+1}$ is defined as:
\begin{equation}
\mathcal{A}^P_t = \pi(\xi(S^P_t))
\end{equation}
\begin{equation}
\mathcal{S}^P_{t+1} = \mathcal{T}(\mathcal{S}^P_t, \mathcal{A}^P_t)
\end{equation}
Given a time $t$, the system can be also represented as a graph $G_t = (V_t, E_t)$, 
  where $E_t$ is built from $\mathcal{N}$.
  Each node is then decorated with the local observation perceived at the time $t$: $o^v_t \in O$. 
  Therefore, this graph can be used for being processed by a \ac{GNN} as done in~\cite{DBLP:conf/corl/TolstayaGPP0R19,tolstaya2020learning,DBLP:conf/icra/GosrichMLPYR022} if we set $f_v = o^v_t$.
\subsection{Motivation}
%The definition of distributed controllers for intelligent systems can be divided into two major approaches: 
% \emph{manual design}, where the application is conceived and described through a programming language, 
% and \emph{automatic design}~\cite{brambilla2013swarm}, which employs machine learning techniques to synthesise distributed controllers. 
%
%Our work lies between these two approaches, specifically in the areas of field-based coordination and \ac{MAARL} with the use of \acp{GNN}.
As mentioned the above, our work falls between automatic and manual approaches, specifically in the areas of field-based coordination and \ac{MAARL} with the use of \acp{GNN}.
%
Compared to earlier works on field-based coordination, 
 our approach builds on the concept of co-fields~\cite{DBLP:journals/pervasive/MameiZL04}, 
 where agents exploit the field as a stigmergic medium to receive system-wide information and apply reasoning to this data.
%  
We present a subsequent approach where agent intelligence is synthesised through \ac{MAARL} and the use of \ac{GNN} to learn a local representation from the neighbourhood.
%
By learning a smart policy directly in the environment, the agent becomes capable of adapting its behaviour to new situations.

@@

As value-based algorithm we relied on \ac{DQN}~\cite{mnih2015playing} %\ga{to introduce in many rl??} 
with two major modifications (see \Cref{alg:dqn}): 
